Contracts for Contracts

### 2023-06-03 paper/introduction.tex
more intro

Indeed, in recent years, we have seen a number of failures of smart contract
systems due to attacks utilizing vulnerable contract programs.
For example, in the famous DAO hack\note{cite}, the unauthorized attacker is
able transfer money by constructing recursive function calls with
the callback mechanism, causing a total lose of \textdollar50 million.
The root cause of this attack is that the deployed contract does not
checks reentrancy defensively before performing critical actions.

@@

To prevent such attacks, the programmer needs to specify those critical
conditions. The execution is allowed when those conditions are met.
However, we observe that existing popular smart contract languages (e.g.
Solidity) do not provide \emph{rich, effective, convenient} means to specify
and enforce contract behaviors.
Simple first-order assertion primitives (e.g., \code{require}) are indeed available,
but they cannot effectively examine callable addresses or function values, which
are a common pattern used in smart contract programs.
Moreover, many of the critical condition in smart contracts are \emph{temporal}
-- the order of happened events matters.
For instance, to defense the DAO attack caused by reentrancy via
external functions, it requires to check that the function of interest cannot
be invoked \emph{before} the current call of the same function is finished.

% encode temporal behavioral checks using ad-hoc lower-level value behaviors

The lack of sufficient linguistic means to specify and enforce subtle but
critical behaviors discourages programmers to use concise higher-level
abstractions when writing smart contract code.
Instead, to be secure, programmers are obliged to write verbose code with
lower-level or no abstractions, which conflates the main business logic with
behavioral checks, further leading to poor readability and maintainability.
Taking the DAO attack as example again, a common way to implement ``reentrancy
guards'' is to manually define an additional boolean state, indicating
if the function is finished.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
In this paper, we argue that \emph{behavioral software
contracts}~\cite{DBLP:conf/tools/Meyer98a} should play a fundamental role in
the
development of \emph{smart contract} programs.  Similar to smart contracts,
behavioral contracts is an agreement too. They specify the assumptions and
guarantees between software components. The specified assumptions and
guarantees will be monitored and reported when violated in languages with
behavioral contracts.

Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
design-by-contract methodology \cite{DBLP:conf/tools/Meyer98a}, various styles
of behavioral contracts have been widely adopted in Java, Haskell, C\#, Python,
Racket, Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \note{cite}
In this paper, we develop a behavioral contract system \lang for the Solidity
language, providing a practical specification and monitoring system
for both higher-order and temporal behaviors.
Solidity is a programming language to build smart contract programs that run on
the Ethereum platform.  As one of the most popular languages for smart
contracts, Solidity has been used \todo{to do cool things}.

>>>>

Indeed, in recent years, we have seen a number of catastrophic failures of
blockchain systems due to attacks to vulnerable contract programs.
For example, in the famous DAO hack \note{cite}, the unauthorized attacker is
able siphon off money by constructing recursive function calls with
the callback mechanism, causing a total lose of \textdollar50 million.
The root cause of this attack is that the deployed contract does not
checks reentrancy defensively before performing critical actions.

@@

To prevent such attacks, the programmer needs to specify those critical
conditions. The execution is allowed only when those conditions are met.
However, we observe that existing popular smart contract languages (e.g.,
Solidity) do not provide \emph{rich, effective, convenient} means to specify
and enforce contract behaviors.
Simple first-order assertion primitives (e.g., \code{require}) are indeed available,
but they cannot effectively examine callable addresses or function values, which
exhibit higher-order behaviors and are a common pattern used in smart contract programs.
Moreover, many of the critical conditions in smart contracts are \emph{temporal}
-- the order of happened events matters.
For instance, to defense the DAO attack caused by reentrancy via
external functions, it requires to check that the function of interest cannot
be invoked \emph{before} the current call of the same function is finished.

The lack of sufficient linguistic means to specify and enforce subtle but
critical behaviors discourages programmers to write clean and maintainable
code with higher-level abstractions.
Instead, programmers are obliged to write verbose low-level code
to implement security checks, which conflates the main business logic with
behavioral checks, further leading to poor readability and maintainability.
The pervasive temporal conditions in distributed blockchain systems further
complicates the situation: these temporal behavioral checks are usually
encoded using value-level checks in ad-hoc manners.
Taking the DAO attack as example again, a common way to implement ``reentrancy
guards'' is to manually define an additional boolean state, indicating
if the function is finished.
And often worse, defensive checks can be easily forgotten or disregarded,
leading to vulnerable code causing real monetary lose.

%Apart from the investigating effort to certify and verify smart contract
%programs, language designers are responsible to empower programmers with better
%linguistic solutions to secure smart contracts in the first place.
To address these issues, we argue that \emph{behavioral software
contracts}~\cite{DBLP:conf/tools/Meyer98a} should play a fundamental role in
the development of \emph{smart contract} programs.
Similar to smart contracts, behavioral contracts is an agreement too. They
specify assumptions and guarantees between software components, just as
smart contracts specify assumptions and guarantees between economic parties.
In languages with behavioral contracts, the specified assumptions and
guarantees will be monitored and reported when violated.

Pioneered by the Eiffel language \cite{DBLP:books/ph/Meyer91} and the
design-by-contract methodology \cite{DBLP:conf/tools/Meyer98a}, various styles
of behavioral contracts have been widely adopted in Java, Haskell, C\#, Python,
Racket, Elixir \cite{DBLP:conf/erlang/0001BBHMEF22}, etc. \note{cite}
In this paper, we develop a behavioral contract system \lang for the Solidity
language, providing a practical specification and monitoring system
for both higher-order and temporal behaviors.
Solidity is a programming language to build smart contract programs that run on
the Ethereum platform.  As one of the most popular ecosystem of smart
contracts, Solidity has been used \todo{to do cool things}.
